---
title: "Change Detection on the Walker Lake"
author: "Christabel Edena Ansah"
date: "April 22, 2019"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
library(raster)
library(rgdal)  
library(RStoolbox)
library(ggplot2) 
library(randomForest)
library(maptools)
library(sp)
library(rgeos)
library(ggmap)
library(ggsn)
library(maps)
setwd("C:/Users/User/Desktop/R Project")
```
## Introduction
  
  + A natural lake, in the Great Basin in western Nevada in the United States.
  
  + 11 mi (17 km) long and 5 mi (8 km) wide.
  
  + Fed from the north by the Walker River and has no natural outlet.
  
  + Over the years changes have occured on the lake and its environs.
  
  + There has been reductions in inflow and continued evaporation of water from the lake.
  
  + These changes were detected using Random Forest and Maximum Likelihood Classification. 
  
## Problem Statement
  
  + The introduction of agriculture upstream of Walker Lake has resulted in the diversion of water from the Walker River and its tributaries for irrigation.
  
  + These diversions have resulted in a severe drop in the level of the lake. 
  
  + The reduction in the volume of water has resulted in a higher concentration of total dissolved solids. 
  
  + Decline of the lake's fishery which is having a dramatic impact on the species of birds using the lake.
  
## Downloaded Landsat 5 and 8 scene for this study
```{r, echo=TRUE, eval=TRUE}
#Load both Landsat 5 and 8 scenes
LS_1984 <- brick("Walker_Lake/1984/1984_Composite/Tif/1984.tif")
LS_2018 <- brick("Walker_Lake/2018/2018_Composite/Tif/2018.tif")
```
```{r, echo=TRUE, eval=TRUE, results='hide', message=FALSE}
#Load the study area boundary shapefile
Study_area <- readOGR("Walker_Lake/Study_Area_Boundary/Boundary.shp")

```
## Reprojecting coordinate system
  + The coordinate system of Landsat scenes was WGS_1984_UTM_Zone_11N. 
  
  + The same was used when creating the study area boundary shapefile in ArcMap.
  
```{r, echo=TRUE,eval=FALSE} 
projection(LS_1984)
projection(LS_2018)
projection(Study_area)
```
```{r, echo=TRUE,eval=TRUE} 
Study_area <- spTransform(Study_area, CRS(proj4string(LS_1984)))
```
```{r, echo=TRUE,eval=FALSE} 

#Set NA value to RASTER NA value.
NAvalue(LS_1984) <- 0
NAvalue(LS_2018) <- 0 
```

## Plotting Study Area
```{r, echo=TRUE,eval=FALSE} 
#Plot the study area on top of the Landsat image to see if the reprojection has worked using the plotRGB tool
plotRGB(LS_1984, r=1, g=4, b=7, stretch="lin")
plotRGB(LS_2018, r=5, g=6, b=4, stretch='lin')
plot(Study_area, col="green", add=TRUE)

#Resize Landsat scenes to the required study area using the crop function
StudyArea_1984 <- crop(LS_1984, Study_area)
StudyArea_2018 <- crop(LS_2018, Study_area)
plotRGB(StudyArea_1984,r=1, g=4, b=7, stretch="lin")
plotRGB(StudyArea_2018,r=5, g=6, b=4, stretch="lin")
```

##  Walker Lake 1984
```{r, echo=FALSE,eval=TRUE}
#Resize Landsat scenes to the required study area using the crop function
StudyArea_1984 <- crop(LS_1984, Study_area)
StudyArea_2018 <- crop(LS_2018, Study_area)
plotRGB(StudyArea_1984,r=1, g=4, b=7, stretch="lin")
```

## Walker Lake 2018
```{r, echo=FALSE,eval=TRUE}
plotRGB(StudyArea_2018,r=5, g=6, b=4, stretch="lin")
```

## Study Area
```{r, echo=TRUE}
img <- stack(StudyArea_1984, StudyArea_2018)
plotRGB(img,r=1, g=4, b=7, stretch="lin")
td <- shapefile("Walker_Lake/Training_Data/Training_Data.shp")
plot(td, col="yellow", add=T)

```

## Image Classification using RF and MLC
```{r, echo=TRUE,results='hide', message=FALSE}
#Image classification of the two landsat scenes
beginCluster()
set.seed(7)
img1984_2018_rf = 
  superClass(img = img, trainData = td, trainPartition = 0.7, 
              responseCol = "Class_Name", nSamples = 500,
              areaWeightedSampling = TRUE,  
              model = "rf", mode = "classification", verbose = FALSE)

## Image Classification using MLC
set.seed(7)
img1984_2018_mlc = 
  superClass(img = img, trainData = td, trainPartition = 0.7, 
              responseCol = "Class_Name", nSamples = 500, 
              areaWeightedSampling = TRUE,  
              model = "mlc", mode = "classification", verbose = FALSE)
endCluster()
```
## Plotting Classification Output of RF
```{r, echo=TRUE,eval=FALSE}
#Random Forest
windows()

ggR(img1984_2018_rf$map, layer = 1, geom_raster = TRUE,forceCat=TRUE)+
  theme(legend.title = element_text(size = 12, face = "bold"))+
  theme(legend.text = element_text(size = 10))+
  scale_fill_manual(values = c("lightsalmon4", "chocolate1", "darkgreen", 
                    "burlywood3", "chartreuse", "blue"),
                    labels=img1984_2018_rf$classMapping$class,
                    name="class name:\n",
                    na.value = "NA",na.translate=FALSE
  )+
  xlab("")+
  ylab("")
north2(img1984_2018_rf$map, x = 0.1, y = 0.8, scale = 0.09, symbol = 1)

```

## RF Output
  + Overall accuracy attained was 98.9%
  
  + kappa value of 0.986
```{r, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
windows()

ggR(img1984_2018_rf$map, layer = 1, geom_raster = TRUE,forceCat=TRUE)+
  theme(legend.title = element_text(size = 12, face = "bold"))+
  theme(legend.text = element_text(size = 10))+
  scale_fill_manual(values = c("lightsalmon4", "chocolate1", "darkgreen", 
                    "burlywood3", "chartreuse", "blue"),
                    labels=img1984_2018_rf$classMapping$class,
                    name="class name:\n",
                    na.value = "NA",na.translate=FALSE
  )+
  xlab("")+
  ylab("")
north2(img1984_2018_rf$map, x = 0.1, y = 0.8, scale = 0.09, symbol = 1)
```

## Plotting Classification Output of MLC
```{r, echo=TRUE,eval=FALSE}
windows()

ggR(img1984_2018_mlc$map, layer = 1, geom_raster = TRUE,forceCat=TRUE)+
  theme(legend.title = element_text(size = 12, face = "bold"))+
  theme(legend.text = element_text(size = 10))+
  scale_fill_manual(values = c("lightsalmon4", "chocolate1", "darkgreen", 
                    "burlywood3", "chartreuse", "blue"),
                    labels=img1984_2018_mlc$classMapping$class,
                    name="class name:\n",
                    na.value = "NA",na.translate=FALSE
  )+
  xlab("")+
  ylab("")
north2(img1984_2018_mlc$map, x = 0.1, y = 0.785, scale = 0.09, symbol = 1)
```

## MLC Output
  + Overall accuracy attained was 95%
  
  + kappa value of 0.926
```{r, echo=FALSE,eval=TRUE,results='hide',message=FALSE}
windows()

ggR(img1984_2018_mlc$map, layer = 1, geom_raster = TRUE,forceCat=TRUE)+
  theme(legend.title = element_text(size = 12, face = "bold"))+
  theme(legend.text = element_text(size = 10))+
  scale_fill_manual(values = c("lightsalmon4", "chocolate1", "darkgreen", 
                    "burlywood3", "chartreuse", "blue"),
                    labels=img1984_2018_mlc$classMapping$class,
                    name="class name:\n",
                    na.value = "NA",na.translate=FALSE
  )+
  xlab("")+
  ylab("")
north2(img1984_2018_mlc$map, x = 0.1, y = 0.785, scale = 0.09, symbol = 1)
```

## Conclusion
  + RF attained the highest accuracy and Kappa value. 
  
  + Water, land and vegetation are very easy classes to identify.
  
  + This can be attributed to the high accuracies in general.
  
  + RF can be said to be the best classifer in this case.
  
  + Put in place measures to prevent Lake Walker's extinction.


## Reference
  + Beutel, M. W., Horne, A. J., Roth, J. C., & Barratt, N. J. (2001). Limnological effects of anthropogenic 
    desiccation of a large, saline lake, Walker Lake, Nevada. In Saline Lakes (pp. 91-105). Springer, Dordrecht.

  + Dickerson, B. R., & Vinyard, G. L. (1999). Effects of high levels of total dissolved solids in Walker Lake, 
    Nevada, on survival and growth of Lahontan cutthroat trout. Transactions of the American Fisheries Society, 
    128(3), 507-515.

  + To access the script: https://github.com/ChristabelAnsah